---
title: "stats506_hw6"
format: 
  html:
    embed-resources: true
execute: 
  error: true
  echo: true
editor: visual
---

github link: <https://github.com/juanitaW12/stats506>

## Load Library

```{r}
suppressPackageStartupMessages(library(dplyr))
library(dplyr)
library(Lahman)
library(DBI)
library(parallel)
library(future)
library(furrr)
library(microbenchmark)
library(ggplot2)
```

## Load Data

```{r}
# Import the SQLite database of the Lahman data
lahman <- dbConnect(RSQLite::SQLite(), "lahman_1871-2022.sqlite")
lahman
```

```{r}
dbListTables(lahman)
```

In this part, I will import the data required to calculate RF, including playerID, teamID, PO, A and InnOuts from the table **fielding**.

```{r}
sql_line <- "SELECT playerID, teamID, PO, A, InnOuts From fielding"
data <- dbGetQuery(lahman, sql_line)
```

```{r}
head(data)
```

As shown above, a single player may play for multiple teams over the course of their career. Next, I need to deal with the missing values and outliers.

## Cleaning Data

### Missing Values

```{r}
# show if any column has missing values
sapply(data, function(x) any(is.na(x)))


# calculate the percentage of missing values
colSums(is.na(data))/nrow(data)

```

Only the InnOuts column contains missing values. Since this column is used as the **denominator** and the proportion of missing values is only 0.2 (20%), I will remove all rows with missing values in this column to ensure data integrity.

### **If Denominator is 0**

```{r}
nrow(data[data$InnOuts == 0, ])/nrow(data)
```

Since the proportion of rows where InnOuts equals zero is less than 0.2% (0.002), I will remove all such rows to ensure the integrity of the analysis.

```{r}
data <- data[data$InnOuts != 0, ]
```

```{r}
# add a new column of RF
data <- data %>%
   mutate(RF = 3 * (PO + A) / InnOuts)
```

## A. Calculating the Average RF for Each Team

In this part, I will use three methods to calculate average RF for each team.

### Without any Parallel Processing

The following function was adapted with assistance from ChatGPT.

```{r}
# generate a function

stratified_bootstrapping <- function(data = data, n_bootstrap = 1000) {
  resampling_rf <- replicate(n_bootstrap, {
    # Group by team and resample
    resampled_data <- data %>%
      group_by(teamID) %>%
      group_modify( ~ sample_n(.x, size = n(), replace = TRUE)) %>%
      ungroup()
    
    # Calculate the team average RF after resampling
    resampled_data %>%
      group_by(teamID) %>%
      summarise(avg_RF_sample = mean(RF, na.rm = TRUE)) %>%
      ungroup()
  }, 
  # Preserve Original Structure
  simplify = FALSE)
  
  # Merge all the resampling results
  boot_results <- do.call(rbind, resampling_rf)
  
  return(boot_results)
}

```

```{r}
# calculate the execution time
start_time_basic <- Sys.time()

# do stratified bootstrapping
set.seed(1000)
bootstrap_results_basic <- stratified_bootstrapping(data, 1000)

end_time_basic <- Sys.time()
execution_time_basic <- end_time_basic - start_time_basic
```

```{r}
# calculate estimated RF and the standard deviation
RF_std_basic <- bootstrap_results_basic %>%
  group_by(teamID) %>%
  summarise(avg_RF_basic = mean(avg_RF_sample), SE_RF_basic = sd(avg_RF_sample))

```

```{r}
head(RF_std_basic)
```

### Parallel Processing

```{r}
# set parallel environment
cl <- makeCluster(detectCores() - 1)

# Load necessary libraries on worker nodes
clusterEvalQ(cl, library(dplyr))

# Export the dataset to worker nodes
clusterExport(cl, "data")

```

```{r}
# do parallel stratified bootstrapping
bootstrap_parallel <- function(data = data, n = 1000, cl = cl){
  parallel_resample_RF <- parLapply(
    cl,
    1:n, # bootstrap iterations
    function(x) {# generate an anonymous function
      # Group by team and resample
      resampled_data <- data %>%
        group_by(teamID) %>%
        group_modify( ~ sample_n(.x, size = n(), replace = TRUE)) %>%
        ungroup()
    
      # Calculate the team average RF after resampling
      resampled_data %>%
        group_by(teamID) %>%
        summarise(avg_RF_sample = mean(RF, na.rm = TRUE)) %>%
        ungroup()
      }
    )
  results_parallel <- do.call(rbind, parallel_resample_RF) 
  
  return(results_parallel)
  
}

```

```{r}
start_time_parallel <- Sys.time()

set.seed(1000)
bootstrap_results_parallel <- bootstrap_parallel(data, 1000, cl)
stopCluster(cl)

end_time_parallel <- Sys.time()
execution_time_parallel <- end_time_parallel - start_time_parallel
```

```{r}
# calculate estimated RF and the standard deviation
RF_std_parallel <- bootstrap_results_parallel %>%
  group_by(teamID) %>%
  summarise(avg_RF_parallel = mean(avg_RF_sample), SE_RF_parallel = sd(avg_RF_sample))
```

```{r}
head(RF_std_parallel)
```

### With Future Package

```{r}
# set parallel plan
plan(multisession)
```

```{r}
# do stratified bootstrapping
bootstrap_future <- function(data = data, n = 1000) {
  results_future <- future_map_dfr(
    1:n,
    # 1000 bootstrap iterations
    ~ { # generate an anonymous function
      # Group by team and resample
      resampled_data <- data %>%
        group_by(teamID) %>%
        group_modify( ~ sample_n(.x, size = n(), replace = TRUE)) %>%
        ungroup()
      # Calculate the team average RF after resampling
      resampled_data %>%
        group_by(teamID) %>%
        summarise(avg_RF_sample  = mean(RF, na.rm = TRUE)) %>%
        ungroup()
      }, 
    .options = furrr_options(seed = TRUE) # Specify parallel-safe seed management
    )
  
  return(results_future)
  }
  
```

```{r}
start_time_future <- Sys.time()

# set random seed
set.seed(1000)
bootstrap_results_future <- bootstrap_future(data, 1000)

end_time_future <- Sys.time()
execution_time_future <- end_time_future - start_time_future
```

```{r}
# calculate estimated RF and the standard deviation
RF_std_future <- bootstrap_results_future %>%
  group_by(teamID) %>%
  summarise(avg_RF_future = mean(avg_RF_sample), SE_RF_future = sd(avg_RF_sample))
```

```{r}
head(RF_std_future)
```

## B. TOP10 Team

The estimated RF is the average RF of all samples.

```{r}
top10_teams_basic <- RF_std_basic %>%
  arrange(desc(avg_RF_basic)) %>%
  slice(1:10)

top10_teams_parallel <- RF_std_parallel %>%
  arrange(desc(avg_RF_parallel)) %>%
  slice(1:10)

top10_teams_future <- RF_std_future %>%
  arrange(desc(avg_RF_future)) %>%
  slice(1:10)

```

```{r}
# Merging data frames based on 'teamID'
compare_results <- merge(
  top10_teams_basic, 
  top10_teams_parallel, 
  by = "teamID", 
  all = FALSE  # Keep only matching rows
)

compare_results <- merge(
  compare_results, 
  top10_teams_future, 
  by = "teamID", 
  all = FALSE  # Keep only matching rows
)
```

```{r}
compare_results
```

I will round the data to improve clarity and then reanalyze it for further investigation.

```{r}
compare_results <- compare_results %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

```{r}
compare_results
```

## C. Performance Difference between the Versions

Firstly, I will compare the performance difference of three versions from the following two aspects.

### Results Comparision

As shown in the compare_results table, the estimated RF values and associated standard errors for the teams with the 10 highest RFs from the three approaches are highly consistent, with virtually no significant differences.

For some teams, such as KEO, the avg_RF values show slight differences at the third decimal place. These differences are likely due to minor variations in random sampling and are statistically negligible.

### Execution Time

```{r}
print(paste("Execution Time of basic method:", round(execution_time_basic,2), "mins"))

print(paste("Execution Time of parallel method:", round(execution_time_parallel,2), "mins"))

print(paste("Execution Time using future:", round(execution_time_future,2), "mins"))
```

As is shown:

1.  The execution time of **Basic Method** is **5.96 minutes**, which is the slowest because it does not leverage multi-core processing. This method runs sequentially on a single thread, processing one task at a time.

2.  The execution time of **Best Use Case** is **1.64 minutes**, which is the quickest because this method distributes tasks across multiple cores, significantly reducing runtime. It can takes advantage of multi-core CPUs to execute tasks concurrently.

3.  The execution time of **Future-Based Method** is **2.92 minutes**, which is the medium running time. This approach also leverages multi-core processing but with more flexibility. Although it is slower than the parallel method but not too much, it is more adaptable to advanced architectures.
